---
title: "PS01 P1"
---






# A. Import the data

```{r}
dat <- read.csv("p1/data.csv", sep = ",", header = TRUE)
head(dat)
```





# B. Verify dataset characteristics

```{r}
# number of observations
nrow(dat)
```

```{r}
# number of distinct cultivars
length(unique(dat$Cultivar))
```

```{r}
# number of distinct seasons
length(unique(dat$Season))
```

```{r}
# Observation per session
season_table <- table(dat$Season)
season_table
```




The dataset has 320 observations, 40 distinct cultivars, 2 seasons.
Season 1 has 160 observations, season 2 has 160 observations.



# C. Answer to following questions

## 1. Correlation between PH and GY

```{r}
cor(dat$PH, dat$GY)
```




The correlation between PH and GY is $0.1233$.



## 2. Correlation by season

```{r}
dat1 <- dat[dat$Season == 1, ]
cor(dat1$PH, dat1$GY)
```

```{r}
dat2 <- dat[dat$Season == 2, ]
cor(dat2$PH, dat2$GY)
```




The correlation between PH and GY is $0.2309$ in season 1, but $-0.0226$ in season 2.



## 3. Grain yield of the observation with the greatest plant height

```{r}
dat[which.max(dat$PH), ]
```




The grain yield of the soybean observation with the greatest plant height is $3304.678$.



## 4. Percentage of observations with MHG above the overall average

```{r}
mean_MHG <- mean(dat$MHG)
mean(dat$MHG > mean_MHG)
```




There are $47.19\%$ of the observation with MHG above the overall average.



## 5. Cultivar with the highest average grain yield

```{r}
mean_GY_by_cultivar <- aggregate(GY ~ Cultivar, data = dat, FUN = mean)
mean_GY_by_cultivar[which.max(mean_GY_by_cultivar$GY), ]
```




The cultivar with highest average grain yield is "98R30 CE", whose average grain yield is $3965$.


# D. Table of average values (overall + by season)

```{r}
var_names <- names(dat)[4:11]
Overall <- sapply(dat[var_names], mean)
Season1 <- sapply(dat1[var_names], mean)
Season2 <- sapply(dat2[var_names], mean)
rbind(Overall, Season1, Season2)
```





# E. T-test comparing grain yield between seasons

## Hypotheses

Denote the mean grain yield of season $i$ as $\mu_i$. Then, the null hypothesis is

$$ H_0: \mu_1 = \mu_2 $$

And the alternative hypothesis is

$$ H_A: \mu_1 \neq \mu_2 $$

First, calculate $t$ using

$$ t = \frac{\bar{\mu}_1 - \bar{\mu}_2}{\sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}} $$

```{r}
mu1 <- mean(dat1$GY)
mu2 <- mean(dat2$GY)
var1 <- var(dat1$GY)
var2 <- var(dat2$GY)
n1 <- length(dat1$GY)
n2 <- length(dat2$GY)
t <- (mu1 - mu2) / sqrt(var1 / n1 + var2 / n2)
t
```




Then we calculate df by

$$ df = \frac{\left(\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}\right)^2}{\frac{(s_1^2/n_1)^2}{n_1-1} + \frac{(s_2^2/n_2)^2}{n_2-1}} $$

```{r}
df <- (var1 / n1 + var2 / n2)^2 / ((var1 / n1)^2 / (n1 - 1) + (var2 / n2)^2 / (n2 - 1))
df
```




Finall, the p value can be calculated by

$$ p = 2 \times (1-F_t(|t|, df)) $$

```{r}
p <- 2 * (1 - pt(abs(t), df))
p
```




Since $p=0.7255 > 0.05$, we cannot reject the null hypothesis. And there's no evidence to suggest that the difference in means is not equal to 0.