---
title: "PS01 P2"
---


# A. Import the data

```{r}
# Read the adult.data file
adult <- read.csv("./p2/adult.data",
                  header = FALSE,
                  sep = ",",
                  strip.white = TRUE,
                  na.strings = "?")
head(adult)
```


# B. Clean and simplify variable names  

We follow the names given in `adult.names`, but change kebab-case to dots.

```{r}
names(adult) <- c(
  "age",
  "workclass",
  "fnlwgt",
  "education",
  "education.num",
  "marital.status",
  "occupation",
  "relationship",
  "race",
  "sex",
  "capital.gain",
  "capital.loss",
  "hours.per.week",
  "native.country",
  "income"
)

head(adult)
```


# C. Examine missing values

```{r}
missing_counts <- colSums(is.na(adult))
missing_counts
```


# D. Restrict sample to individuals working at least 10 hours

```{r}
n_before <- nrow(adult)

subset_d <- adult[adult$hours.per.week >= 10, ]

n_after <- nrow(subset_d)

c(n_before, n_after)
```


Before restrict the sample, there are $32,561$ observations. And after the restriction, there are $32,103$ observations.

# E. Remove observations missing occupation, workclass, or native.country  

```{r}
subset_e <- subset_d[!is.na(subset_d$occupation) & !is.na(subset_d$workclass) & !is.na(subset_d$native.country), ]
nrow(subset_e)
```


After removing observations with missing values on these columns, there are $29,873$ observations.
Occupation, workclass, or native country are important variables for income.
Missing values in these variables may affect the analysis.

# F. Restrict observations further more

```{r}
subset_f <- subset_e[subset_e$age >= 18 & subset_e$age <= 90 & subset_e$education != "Preschool", ]
nrow(subset_f)
```



# G. Report final sample size

There are $29,523$ observations after the restriction.