---
title: "PS01"
format:
  html:
    embed-resources: true
editor: visual
---

GitHub repository: <https://github.com/suyi-afk/stats506>

# Problem1

# A. Import the data

```{r}
dat <- read.csv("p1/data.csv", sep = ",", header = TRUE)
head(dat)
```

# B. Verify dataset characteristics

```{r}
# number of observations
nrow(dat)
```

```{r}
# number of distinct cultivars
length(unique(dat$Cultivar))
```

```{r}
# number of distinct seasons
length(unique(dat$Season))
```

```{r}
# Observation per session
season_table <- table(dat$Season)
season_table
```

The dataset has 320 observations, 40 distinct cultivars, 2 seasons. Season 1 has 160 observations, season 2 has 160 observations.

# C. Answer to following questions

## 1. Correlation between PH and GY

```{r}
cor(dat$PH, dat$GY)
```

The correlation between PH and GY is $0.1233$.

## 2. Correlation by season

```{r}
dat1 <- dat[dat$Season == 1, ]
cor(dat1$PH, dat1$GY)
```

```{r}
dat2 <- dat[dat$Season == 2, ]
cor(dat2$PH, dat2$GY)
```

The correlation between PH and GY is $0.2309$ in season 1, but $-0.0226$ in season 2.

## 3. Grain yield of the observation with the greatest plant height

```{r}
dat[which.max(dat$PH), ]
```

The grain yield of the soybean observation with the greatest plant height is $3304.678$.

## 4. Percentage of observations with MHG above the overall average

```{r}
mean_MHG <- mean(dat$MHG)
mean(dat$MHG > mean_MHG)
```

There are $47.19\%$ of the observation with MHG above the overall average.

## 5. Cultivar with the highest average grain yield

```{r}
mean_GY_by_cultivar <- aggregate(GY ~ Cultivar, data = dat, FUN = mean)
mean_GY_by_cultivar[which.max(mean_GY_by_cultivar$GY), ]
```

The cultivar with highest average grain yield is "98R30 CE", whose average grain yield is $3965$.

# D. Table of average values (overall + by season)

```{r}
var_names <- names(dat)[4:11]
Overall <- sapply(dat[var_names], mean)
Season1 <- sapply(dat1[var_names], mean)
Season2 <- sapply(dat2[var_names], mean)
rbind(Overall, Season1, Season2)
```

# E. T-test comparing grain yield between seasons

## Hypotheses

Denote the mean grain yield of season $i$ as $\mu_i$. Then, the null hypothesis is

$$ H_0: \mu_1 = \mu_2 $$

And the alternative hypothesis is

$$ H_A: \mu_1 \neq \mu_2 $$

First, calculate $t$ using

$$ t = \frac{\bar{\mu}_1 - \bar{\mu}_2}{\sqrt{\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}}} $$

```{r}
mu1 <- mean(dat1$GY)
mu2 <- mean(dat2$GY)
var1 <- var(dat1$GY)
var2 <- var(dat2$GY)
n1 <- length(dat1$GY)
n2 <- length(dat2$GY)
t <- (mu1 - mu2) / sqrt(var1 / n1 + var2 / n2)
t
```

Then we calculate df by

$$ df = \frac{\left(\frac{s_1^2}{n_1} + \frac{s_2^2}{n_2}\right)^2}{\frac{(s_1^2/n_1)^2}{n_1-1} + \frac{(s_2^2/n_2)^2}{n_2-1}} $$

```{r}
df <- (var1 / n1 + var2 / n2)^2 / ((var1 / n1)^2 / (n1 - 1) + (var2 / n2)^2 / (n2 - 1))
df
```

Finall, the p value can be calculated by

$$ p = 2 \times (1-F_t(|t|, df)) $$

```{r}
p <- 2 * (1 - pt(abs(t), df))
p
```

Since $p=0.7255 > 0.05$, we cannot reject the null hypothesis. And there's no evidence to suggest that the difference in means is not equal to 0.

# Problem 2

# A. Import the data

```{r}
# Read the adult.data file
adult <- read.csv("./p2/adult.data",
                  header = FALSE,
                  sep = ",",
                  strip.white = TRUE,
                  na.strings = "?")
head(adult)
```

# B. Clean and simplify variable names

We follow the names given in `adult.names`, but change kebab-case to dots.

```{r}
names(adult) <- c(
  "age",
  "workclass",
  "fnlwgt",
  "education",
  "education.num",
  "marital.status",
  "occupation",
  "relationship",
  "race",
  "sex",
  "capital.gain",
  "capital.loss",
  "hours.per.week",
  "native.country",
  "income"
)

head(adult)
```

# C. Examine missing values

```{r}
missing_counts <- colSums(is.na(adult))
missing_counts
```

# D. Restrict sample to individuals working at least 10 hours

```{r}
n_before <- nrow(adult)

subset_d <- adult[adult$hours.per.week >= 10, ]

n_after <- nrow(subset_d)

c(n_before, n_after)
```

Before restrict the sample, there are $32,561$ observations. And after the restriction, there are $32,103$ observations.

# E. Remove observations missing occupation, workclass, or native.country

```{r}
subset_e <- subset_d[!is.na(subset_d$occupation) & !is.na(subset_d$workclass) & !is.na(subset_d$native.country), ]
nrow(subset_e)
```

After removing observations with missing values on these columns, there are $29,873$ observations. Occupation, workclass, or native country are important variables for income. Missing values in these variables may affect the analysis.

# F. Restrict observations further more

```{r}
subset_f <- subset_e[subset_e$age >= 18 & subset_e$age <= 90 & subset_e$education != "Preschool", ]
nrow(subset_f)
```

# G. Report final sample size

There are $29,523$ observations after the restriction.

# Problem 3

The function `nextHappy` computes the next number in a happy-number sequence.

```{r}
#' Compute the next number in a happy-number sequence
#'
#' Takes a positive integer n, splits it into digits, squares each digit, and returns the sum of those squares.
#'
#' @param n numeric. The number whose digit-square-sum is computed.
#' @return numeric. The sum of squares of the digits of n.
#' @examples
#' nextHappy(1) # 1
#' nextHappy(12)  # 1^2 + 2^2 = 1 + 4 = 5
nextHappy <- function (n) {
    sum(as.numeric(strsplit(as.character(n), '')[[1]])^2)
}
```

The function `happySequence` test if given number is happy. And output the sequence of calculating happy numbers if it is happy.

```{r}
#' Generate the happy-number sequence for an integer
#' 
#' @param n numeric. A positive integer for the start of sequence.
#' @return list containing two elements:
#'     $sequence: the sequence starts with n and ends with 1, only presented when n is happy.
#'     $isHappy: logical. Whether n is happy.
happySequence <- function (n) {
    #' A helper recursive function to generate the happy number sequence.
    #' Returns empty list when the number is unhappy.
    genSeq <- function (n, prev=c()) {
        succ = c(prev, n)
        if (n == 1) {
            succ
        } else if (n %in% prev) {
            c()
        } else {
            genSeq(nextHappy(n), succ)
        }
    }
    sequence = genSeq(n)
    if (length(sequence) == 0) {
        list(isHappy = FALSE)
    } else {
        list(sequence = sequence, isHappy = TRUE)
    }
}
```

With the help of above functions. We working on numbers 1 to 1000.

```{r}
happyList <- sapply(seq(1, 1000), FUN=function (n) {
    happySequence(n)$isHappy
})
lengthList <- sapply(seq(1, 1000), FUN=function (n) {
    sequence <- happySequence(n)$sequence
    if (happySequence(n)$isHappy) {
        length(sequence)
    } else {
        NA
    }
})
c(sum(happyList), mean(happyList), which.max(lengthList))
```

```{r}
happySequence(356)
```

There are 143 (14.3%) happy numbers in 1 to 1000. And number 356 has the longest sequence, which contains 7 numbers: 356, 70, 49, 97, 130, 10, 1.
