# Problem 1: Random Walk with Momentum

## a. Simulation function

Before moving on, let's create a function that helps sanitize the input of momentum walk.

We need to verify that $n$ is positive integer, and $p$ is a probability in $[0, 1]$.

```{r}
#' A common helper function which check if
#'     * n is positive integer, and
#'     * p in [0, 1]
#' 
#' @param f the implementation of momentum walk
#' @return implementation with parameter guarded
momentum_walk_santilize <- function (f) {
    function (n, p) {
        if (!is.numeric(n) || (n < 1) || as.integer(n) != n) {
            stop("n must be a positive integer")
        }
        if (!is.numeric(p) || p < 0 || p > 1) {
            stop("p must be a probability between 0 and 1")
        }
        f(n, p)
    }
}
```

Then, let's focus on the first implementation.

The first implementation using vectorized apporach. 1. We use `runif > p` to check if we need a new direction. 2. Calculate the accumulated sum, so we know how many directions we need to decide, and in each step which direction we need to use. 3. Use $\mathrm{i}^{d}$ to get 4 directions, and sum these complex numbers to get the destination. 4. Convert the destination to two numbers

```{r}
#' Consider a particle moving in two dimensions. The particle starts at
#' $$ (x,y)=(0,0) $$
#' At each step, the particle moves one unit in one of four directions:
#' right:(+1,0), left:(-1,0), up:(0,+1), down:(0,−1)
#' For the first step, choose one of the four directions with equal
#' probability (which means, 1/4).
#' For every subsequent steps, With probability p, the particle moves
#' in the same direction as the previous step.
#' With probability 1−p, it chooses one of the four directions at
#' random, with equal probability.
#' 
#' @param n number of steps
#' @param p probability of choosing the same direction as the previous step
#'          not that for probability 1-p, there are still chances of choosing
#'          the same direction as the previous step
#' @return two integers representing the destination of moving
momentum_walk <- momentum_walk_santilize(function (n, p) {
    has_turn <- c(TRUE, (runif(n - 1) > p))
    turn_count <- cumsum(has_turn)
    total_turn <- turn_count[length(turn_count)]
    directions <- sample(1:4, total_turn, replace = TRUE)[turn_count]
    destination = sum(1i ** directions)
    c(Re(destination), Im(destination))
})
```

## b. Implement the simulation

Here is the loop approach.

1.  Repeat until no remaining steps needed, starting from `n` steps
2.  Since the continuous steps in single direction follow Geometric(1-p), we get the number of steps in this direction directlty using `rgeom`.
3.  Then, we sum all the steps to get destination

```{r}
#' Same as `momentum_walk` but with loops
#' 
#' @param n number of steps
#' @param p probability of keeping previous direction
#' @return the destination
momentum_walk1 <- momentum_walk_santilize(function(n, p) {
    if (p == 1) {
        direction <- sample(1:4, size=1)
        destination <- 1i ** direction * n
        return(c(Re(destination), Im(destination)))
    }

    steps_remain <- n
    destination <- 0
    while (steps_remain > 0) {
        steps <- min(steps_remain, rgeom(1, 1 - p) + 1)
        direction <- sample(1:4, size=1)
        steps_remain <- steps_remain - steps
        destination <- destination + 1i ** direction * steps
    }
    c(Re(destination), Im(destination))
})
```

Here is the `apply` approach.

1.  Chose a direction initially.
2.  Repeat `n` times, each time choosing a new direction with probability $1-p$, or keep the previous direction otherwise.
3.  Get a list of each steps, and then sum them up get destination.

```{r}
#' Same as `momentum_walk` but with applies
#' 
#' @param n the number of steps
#' @param p the probability of keeping previous direction
#' @return the destination
momentum_walk2 <- momentum_walk_santilize(function (n, p) {
    direction <- sample(1:4, size=1)
    destination <- sum(sapply(1:n, function (step) {
        direction <<- if (runif(1) > p) sample(1:4, size=1) else direction
        1i ** direction
    }))
    c(Re(destination), Im(destination))
})
```

We can test our code

```{r}
print(momentum_walk(10, 0.6))
print(momentum_walk(1000, 0.6))
print(momentum_walk1(10, 0.6))
print(momentum_walk1(1000, 0.6))
print(momentum_walk2(10, 0.6))
print(momentum_walk2(1000, 0.6))
```

## c. Reproducibility

In order to ensure the reproducibility, we need to config random seed in our code.

```{r}
set.seed(42)
print(momentum_walk(10, 0.6))
set.seed(42)
print(momentum_walk(10, 0.6))

set.seed(42)
print(momentum_walk1(10, 0.6))
set.seed(42)
print(momentum_walk1(10, 0.6))

set.seed(42)
print(momentum_walk2(10, 0.6))
set.seed(42)
print(momentum_walk2(10, 0.6))
```

```{r}
set.seed(42)
print(momentum_walk(1000, 0.6))
set.seed(42)
print(momentum_walk(1000, 0.6))

set.seed(42)
print(momentum_walk1(1000, 0.6))
set.seed(42)
print(momentum_walk1(1000, 0.6))

set.seed(42)
print(momentum_walk2(1000, 0.6))
set.seed(42)
print(momentum_walk2(1000, 0.6))
```

Above execution shows that, when seed is given, the output is reproducible.

## d. Does momentum matter?

```{r}
as.data.frame(t(sapply(c(0, 0.5, 0.9), function (p) {
    set.seed(42)
    destination = momentum_walk(1000, p)
    c(p = p, destination = destination, distance = sqrt(sum(destination ^ 2)))
})))
```

From these three simulations, there is no clear pattern that the distance increases as $p$ increases. Although higher momentum should generally lead to a larger distance, a single simulation for each value of $p$ is not enough to show this pattern clearly because of randomness.

## e. Monte Carlo simulation

So we try to simulate 1,000 times. And see if we can get a better understanding of the distance.

```{r}
simulation_result <- as.data.frame(t(sapply(c(0, 0.25, 0.5, 0.75, 0.9), function (p) {
    set.seed(42)
    mean_distance = mean(sapply(1:1000, function (i) sqrt(sum(momentum_walk(1000, p) ^ 2))))
    c(p = p, mean_distance = mean_distance)
})))

simulation_result
```

```{r}
plot(mean_distance ~ p, data = simulation_result, type = 'b', xlab = 'p', ylab = 'mean distance', main = 'Distance of random walks with momentum')
```

After 1,000 simulations, the pattern is clearer. The mean final distance increases as $p$ increases.

## f. Compare computational performance

We run benchmarks with `n = 1000, p = 0.6`, and `n = 100000, p = 0.6` using all 3 functions.

```{r}
library(microbenchmark)
```

```{r}
benchmark1k <- microbenchmark(
    momentum_walk  = momentum_walk(1000, 0.6),
    momentum_walk1 = momentum_walk1(1000, 0.6),
    momentum_walk2 = momentum_walk2(1000, 0.6),
    times = 100
)

summary(benchmark1k)
```

```{r}
benchmark100k <- microbenchmark(
    momentum_walk  = momentum_walk(100000, 0.6),
    momentum_walk1 = momentum_walk1(100000, 0.6),
    momentum_walk2 = momentum_walk2(100000, 0.6),
    times = 100
)

summary(benchmark100k)
```

Since the code is $\mathcal{O}(n)$ time complexity. It wouldn't be amazing that whin larger $n$, the code consumes more time.

The vectorlized code runs fastest. That's because the vectorized version take the advantage of using underlying C code to boost the performance.

And the later 2 versions cannot be vectorlized, and is \~40x slower.

However, version 1 is still faster, that due to we optimized the code using `rgeom` to move multiple steps at once.

The performance difference becomes larger as $n$ increases.
