# Problem 3: Linear Regression

We first download the file once, so we do not need to download it every time we run the code.

```{r}
youtube.url = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv"
if (!dir.exists('p3')) dir.create('p3')
if (!file.exists('p3/youtube.csv')) download.file(youtube.url, 'p3/youtube.csv')
```

```{r}
youtube <- read.csv('p3/youtube.csv')
```

## a. Data preparation and de-identification

Then, we de-indentify the data by removing variables such as the brand, title, URLs, YouTube channel, and publication information.

```{r}
youtube$brand <- NULL
youtube$superbowl_ads_dot_com_url <- NULL
youtube$youtube_url <- NULL
youtube$kind <- NULL
youtube$id <- NULL
youtube$etag <- NULL
youtube$title <- NULL
youtube$description <- NULL
youtube$thumbnail <- NULL
youtube$channel_title <- NULL
youtube$published_at <- NULL
```



```{r}
dim(youtube)
head(youtube)
```

# b. Exploring potential outcomes

```{r}
summary(youtube[, c("view_count", "like_count", "dislike_count", "favorite_count", "comment_count")])
```

```{r}
par(mfrow = c(1, 5))

hist(youtube$view_count, main = "Views", xlab = "")
hist(youtube$like_count, main = "Likes", xlab = "")
hist(youtube$dislike_count, main = "Dislikes", xlab = "")
hist(youtube$favorite_count, main = "Favorites", xlab = "")
hist(youtube$comment_count, main = "Comments", xlab = "")

par(mfrow = c(1, 1))
```

Since all `favorite_count` are `0`. This column is not suitable for further analysis.

For all other columns, they are heavily right skewed. And take `log1p` before analysis would be a better idea.

```{r}
par(mfrow = c(1, 4))

hist(log1p(youtube$view_count), main = "Log1p(Views)", xlab = "")
hist(log1p(youtube$like_count), main = "Log1p(Likes)", xlab = "")
hist(log1p(youtube$dislike_count), main = "Log1p(Dislikes)", xlab = "")
hist(log1p(youtube$comment_count), main = "Log1p(Comments)", xlab = "")

par(mfrow = c(1, 1))
```

After transforming using `log1p`, the data have a better shape then. And we will use the transformed data in following regressions.

## c. Fitting linear regression models

```{r}
outcomes <- c("view_count", "like_count", "dislike_count", "comment_count")
predictors <- c("funny", "show_product_quickly", "patriotic", "celebrity", "danger", "animals", "use_sex")
```

```{r}
outcome <- outcomes[1]
predictor <- predictors[1]
lm_results <- do.call(rbind, lapply(outcomes, function (outcome) {
    data.frame(t(sapply(predictors, function (predictor) {
        model <- lm(as.formula(paste0("log1p(", outcome, ") ~ as.numeric(", predictor, ") + year")), data=youtube)
        c(outcome=outcome, predictor=predictor, coefficients=summary(model)$coefficients[2, 1], p.value=summary(model)$coefficients[2, 4])
    })))
}))
rownames(lm_results) <- 1:nrow(lm_results)
lm_results$coefficients <- as.numeric(as.character(lm_results$coefficients))
lm_results$p.value <- as.numeric(as.character(lm_results$p.value))
lm_results
```

```{r}
lm_results[lm_results$p.value < 0.05, ]
```

at the $\alpha=0.05$ significance level, there are no pairs of predictors and outcomes that are statistically significantly associated with each other.

```{r}
view_count_results <- lm_results[lm_results$outcome == 'view_count', ]
view_count_results[order(view_count_results$coefficients, decreasing = TRUE), ]
```

For `view_count`, `danger` has the largest positive estimated association with it.

## d. Estimating regression coefficients using matrix algebra

We can calculate the regression
$$ \mathrm{log1p}(ViewCount) \sim Danger + Year $$
using the `lm` function in R as follow

```{r}
summary(lm(log1p(view_count) ~ danger + year, data = youtube))
```

And on the other hand. We try to calculate it manually using formula

$$ \hat{\mathbf{\beta}} = \left(\mathbf{X}^\top \mathbf{X}\right)^{-1}\mathbf{X}^\top \mathbf{y} $$

```{r}
present_data <- na.omit(youtube[, c("view_count", "danger", "year")])

y <- log1p(present_data$view_count)

X <- cbind(
  "(Intercept)" = 1, 
  danger = as.numeric(present_data$danger), 
  year = present_data$year
)

beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y
beta_hat
```

And we can see that they have the same coefficients.